---
title: "Tunis demo Friday"
author: "Daniel"
date: "11/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = 'figures/')
library(edgeR)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
```

```{r read-data}
# Read the counts table
counts <- read_tsv(
  '../data/atacama-soils.asvtable.tsv',
  col_types = cols(
    .default = col_integer(), seqid = col_character()
  )
) %>%
  # Make the "wide" table "long"
  gather(sample, count, 2:67) %>%
  # Get rid of rows with zero counts
  filter(count > 0) %>%
  group_by(sample) %>%
  filter(sum(count) >= 1000) %>%
  ungroup() %>%
  group_by(sample) %>%
  mutate(relab = count/sum(count)) %>%
  ungroup()

# Read the samples table
samples <- read_tsv(
  '../data/sample_metadata_single_header.tsv',
  col_types = cols(
    .default = col_double(),
    SampleID = col_character(),
    BarcodeSequence = col_character(),
    LinkerPrimerSequence = col_character(),
    Elevation = col_integer(),
    ExtractGroupNo = col_character(),
    TransectName = col_character(),
    SiteName = col_character(),
    Depth = col_integer(),
    TOC = col_integer(),
    Vegetation = col_character(),
    Description = col_character()
  )
)

# Read the taxonomy
taxonomy <- read_tsv(
  '../data/taxonomy.tsv',
  col_types = cols(.default = col_character(), Confidence = col_double())
) %>%
  # Get rid of the "D_[0-9]__" string that starts each individual taxon
  mutate(
    Taxon = gsub('D_[0-9]__', '', Taxon)
  ) %>%
  # Separate the long taxon string into separate taxa
  separate(
    Taxon,
    c(
      'domain', 'phylum', 'class', 
      'order', 'family', 'genus', 'species'
    ),
    sep = ';',
    fill = 'right'
  ) %>%
  # Rename the Feature ID so it's seqid, like in the counts table
  rename(seqid = `Feature ID`)
```

```{r topphyla}
taxonomy <- taxonomy %>%
  left_join(
    counts %>%
      inner_join(taxonomy, by = 'seqid') %>%
      group_by(phylum, sample) %>%
      summarise(relab = sum(relab)) %>%
      summarise(mean_relab = mean(relab)) %>%
      ungroup() %>%
      top_n(10, mean_relab) %>%
      transmute(phylum, topphylum = phylum),
    by = 'phylum'
  ) %>%
  replace_na(list(topphylum = 'Other phyla'))
```

# Downsampling the datasets

I'm selecting observations only from the Yungay transect with humidities > 85 and < 50%.

```{r}
d <- samples %>%
  inner_join(counts, by = c('SampleID' = 'sample')) %>% 
  filter(TransectName == 'Baquedano') %>% 
  filter(AverageSoilRelativeHumidity > 85 | AverageSoilRelativeHumidity < 50) %>% 
  mutate(h = ifelse(AverageSoilRelativeHumidity > 75, 'high', 'low')) %>%
  select(SampleID, seqid, count, h) %>%
  arrange(h, SampleID) %>%
  select(-h) %>%
  spread(SampleID, count, fill = 0) %>%
  tibble::column_to_rownames('seqid') 
dge <- d %>%
  DGEList(group = c(1,1,1,1,1,0,0,0,0,0)) %>%
  calcNormFactors() %>%
  estimateDisp()
et <- exactTest(dge)
tt <- topTags(et, n = 100000)
```

